SPERF a scalability profiler for shared-memory parallel applications

### 2022-08-15 thesis/conclusion.tex
organazing the chapters and trying to connect

\section{Contributions*} \label{sec:contributions}

\section{Conclusion Models} \label{sec:conclusion_models}
This paper proposes an energy model based on the operating frequency and the number of cores for a shared memory system. This model serves as a reference for DVFS and DPM optimization problems.

@@

%Future work will demonstrate all possible analyzes achievable with the proposed equation as more advanced DVFS models can be derived from it. For instance, identifying the different phases of a target program will allow more subtle changes in frequency and, perhaps, in the number of active cores to improve further the results presented here.

This work proposes an energy model based on the frequency and number of cores for a full shared memory system. This model can serve as a base for DVFS and DPM optimization problems that include both frequency and active cores, as well for analysis of the contribution of each parameter (ex: parallelism level) to the energy consumption.

@@

\subsection{Conclusions and Future Works Pascal} \label{sec:conclusions_pascal}

The tool has a low level of intrusion added to the program's performance measurement under analysis, which is a fundamental aspect of understanding the program's behavior and scalability capacity. 
Although it does not offer graphic elements to visualize the collected metrics, the tool itself makes it possible to examine the data from the command terminal. It also allows this data to be analyzed by an interface provided in Python or through .json files. 
%Other tools, such as PaScal Viewer, can compensate for the absence of these graphic resources. Indeed, PaScal Viewer is a web-based visualization tool focused on analyzing the scalability trend of parallel applications, which has been integrated with PaScal Analyzer in a complementary way.

\subsection{Conclusion fingerprint tool}

%\cite{Processors2012, Gregg2017}
From the results, we observe that the API present overhead similar or lower to other low-level APIs, with the advantage of being in high abstraction and simplified configuration with a few lines of code, is possible to configure and gather counter data.

%\subsection{FUTURE WORK}
%
%We pretend to use this tool to create a data set of applications behavior and automatically classify any program as belonging to a cluster or a set of clusters. The idea is to have a set of clusters that can describe most applications in this way we can know specific behaviors of the applications. This can be applied to various areas such as voltage scaling and frequency of the processor, once knowing the behavior of a particular program we can have an idea of what is the best strategy to control the frequency in order to save energy or increase performance.

\section{Conclusion and future work phases} \label{sec:conclusion}
The main conclusion of this work was that in the HPC environment, many phases are not necessary to reach optimal energy consumption. However, an average of 35 phases was enough for more than three benchmarks covering the most varied HPC applications.

It was also possible to notice that there is still a lot to improve in optimizing the DVFS algorithms since we obtained an average of 38\% energy savings compared to Ondemand on Linux.

>>>>

\section{Conclusion}

\subsection{Pascal Suite} \label{sec:conclusions_pascal}

The tool has a low level of intrusion added to the program's performance measurement under analysis, which is a fundamental aspect of understanding the program's behavior and scalability capacity. 
Although it does not offer graphic elements to visualize the collected metrics, the tool itself makes it possible to examine the data from the command terminal. It also allows this data to be analyzed by an interface provided in Python or through .json files.

\subsubsection{fingerprint tool}

From the results, we observe that the API present overhead similar or lower to other low-level APIs, with the advantage of being in high abstraction and simplified configuration with a few lines of code, is possible to configure and gather counter data.

\subsection{Application model} \label{sec:conclusion_models}
The proposed energy model based on the operating frequency and the number of cores for a shared memory system can serve as a reference for DVFS and DPM optimization problems.

@@

The proposed energy model based on the frequency and number of cores for a full shared memory system can serve as a base for DVFS and DPM optimization problems that include both frequency and active cores. As well for analysis of the contribution of each parameter (ex: parallelism level) to the energy consumption.

@@

\subsection{Application-phase heuristic} \label{sec:phases_conclusion}
The main conclusion of this work was that in the HPC environment, in addition to the advantages of finding the optimal phase divisions, not many phases are needed to achieve the optimal energy consumption. Moreover, an average maximum of 35 phases was sufficient for the three benchmarks used, covering the most varied HPC applications.
It was also possible to notice that there is still a lot to improve in optimizing the DVFS algorithms since we obtained an average of 38\% energy savings compared to Ondemand on Linux.
We also noticed that there is a relation between the application behavior and the phases locations, independently of their number.

### 2022-08-15 thesis/intro/intro.tex
change subsection to section and labels
samuel suggestions
organazing the chapters and trying to connect

\section{Objectives}

\section{Contributions}

Although much work has been done on DVFS, the focus is still on the consumer electronics and laptop markets. 
For HPC, the notion of energy perception is relatively new~\cite{Feng2003MakingSupercomputing}. 
Moreover, the operational characteristics of non-HPC and HPC systems are significantly different. 
First, the workload on non-HPC systems is very interactive with the end-user, but the workload on the HPC platform is not.
Second, activities conducted on a non-HPC platform tend to share more machine resources.
In contrast, in HPC, each job often runs with dedicated resources. 
Third, an HPC system is usually much larger than a non-HPC system, making it more challenging to gather information, organize, and execute global decisions.
Therefore, it is worthwhile to investigate whether a DVFS scheduling algorithm, which works well for conventional computing, remains effective for HPC.

Our paper proposes a full-system energy model based on the CPU frequency and the number of cores. The model aims to understand and optimize the energy behavior of parallel applications in HPC systems according to application parameters, such as the degree of parallelism and CPU parameters related to dynamic and static power. The proposed model differs from existing ones, including the frequency and number of cores in the same equation for estimating the energy for a specific application in a given configuration. This model can serve as a base for considering DVFS and DPM optimization problems, including frequency and active cores. It can also be used to analyze the contribution of each parameter (ex: level of parallelism) to energy consumption. Furthermore, the number of cores is essential in HPC since applications are designed to run on multiple cores.

The proposed energy model is the product of an application-agnostic power model and an architecture-specific application performance model. The power model is based on the CMOS logic gates power draw as a function of the frequency~\cite{Sarwar1997CmosCalculation, Butzen2007LeakageGates} augmented to include the number of cores. The performance model is based on Amdahl's law \cite{Amdahl1967ValidityCapabilities, Eyerman2010ModelingDesign, Gustafson1988ReevaluatingLaw}, which can be used to estimate runtime in multi-core systems. In addition, this model has been extended to include execution frequency and input size, characterizing the application on the target architecture.

\cref{tab:related_work} summarizes the models comparing the system dependencies and the controllable variables.
\begin{table}[H]
	\footnotesize
	\caption{Related work summary.\label{tab1}}
	\setlength{\tabcolsep}{2.3mm}\begin{tabular}{cccc}
		\toprule
		%\begin{adjustwidth}{-\extralength}{0 cm}

		\textbf{Model}  & \textbf{System Dependency}       & \textbf{Variable}                     & \textbf{Controllable Variables}         \\ \midrule
		Merkel et al. \cite{Merkel2006BalancingSystems} & performance counters    & number of activities & - \\
		Roy et al. \cite{Roy2013AnAlgorithms}&performance counters &io operations, total time    & - \\
		Yakun et al. \cite{Shao2013EnergyProcessor}&number of instructions &frequency&frequency \\
		Lewis et al. \cite{Lewis2008Run-timeSystems}&energy of subcomponents&energy of subcomponents & - \\
		Mills et al. \cite{Mills2014EnergySystems}&power of subcomponents&total time, frequency & frequency \\ 
		Our model & - & frequency, cores, input size & frequency, cores, input size\\ \bottomrule
	\end{tabular}
	%\caption{Related work summary}
	\label{tab:related_work}
	%\bottomrule
\end{table}

The main contributions of the proposed model are:

% Contribuições por capitulo...
\begin{itemize}
	\item Simple model: faster to fit and compute, good for DVFS and DPM optimization.
	\item Parameters with logical meaning: helps to understand the contribution of each specific term.
	\item Analytical analysis: several analyses can be derived from the equation.
	\item Controllable variables: the equation is in the function of parameters that we can control directly.
\end{itemize}

@@

>>>>

\section{Objectives} \label{sec:objectives}
This thesis aims to propose solutions that use the full potential of the systems involved, based on knowledge of the architectures and the applications, as follow:

Develop a framework that helps and automates the process of modeling applications and architecture without interfering with the application.

Propose an analytical energy model for energy consumption in high-performance computing systems applications that both optimize energy consumption in these systems and perform deeper analysis using analytical equations.

Propose an heuristic algorithm, taking into account changes in the behavior of applications and the systems they run.

\section{Contributions} \label{sec:contributions}

In the \cref{chapter:pascal_suite}, we present a tool for automatically measuring and comparing several executions of a parallel application according to various scenarios characterized by the input arrangements, the number of threads, number of cores, and frequencies. Unlike other existing performance analysis tools, the proposed work covers some gaps in specialized features necessary to better understand computational resources scalability trends across configurations while keeping minimal intrusion (less than 1\% overhead).

The \cref{chapter:models}, proposes a full-system energy model based on the CPU frequency and the number of cores. The model aims to understand and optimize the energy behavior of parallel applications in HPC systems according to application parameters, such as the degree of parallelism and CPU parameters related to dynamic and static power. The proposed model differs from existing ones, including the frequency and number of cores in the same equation for estimating the energy for a specific application in a given configuration. This model can serve as a base for DVFS and DPM optimization problems. Including frequency and active cores it can also be used to analyze the contribution of each parameter (ex: level of parallelism) to energy consumption.

\cref{chapter:phases} proposes a methodology that combines measurement data with a heuristic to provide insights into choosing the best phase divisions. Our heuristic can reduce the scan space from $10^{7000}$ to $10^2$ with an average error of 10\% and up to 38\% reduction in energy consumption using optimal distribution compared to standard Linux DVFS. Moreover, we evaluated the trade-off of having too many divisions and the overhead caused. Finally, we have identified that there is usually a limit to the number of phases a running application can benefit from, giving a lower limit to the minimum number of phases.

@@

In the \cref{chapter:theoretical_background} points that are common to the whole thesis are clarified, such as definitions, configurations and benchmarks used. In the next chapter \cref{chapter:pascal_suite} the PascalSuite framework is presented in detail and validated. The energy model is proposed in sequence in \cref{chapter:models}. Finally a heuristic algorithm is presented in \cref{chapter:phases} followed by the conclusion taken from this thesis in \cref{chapter:conclusions}.

### 2022-08-15 thesis/main.tex
Updates from Overleaf
organazing the chapters and trying to connect

	\begin{umonsThesisAbstract}
		Energy consumption is a key to enable exascale High-performance Computing (HPC). Energy-optimized hardware and software combinations still could be inefficient if software operates poorly. 
		Software operation relies on dynamic scaling of frequency and voltage (DVFS) as well as dynamic power management (DPM), but none have priority information on the application, which leads to inefficient software operation. This work proposes an analytical model that can be used to estimate energy-optimal software operation. 
		This work proposes a new energy model that can be used as a basis to optimize DVFS and DPM, as well as to assess the contribution of parameters such as parallelism level and dynamic power to the energy consumption.
		Since in HPC, applications can run on multiple cores and associated frequencies, both of which can be controlled for most systems, the novelty of the model includes such parameters as well as the workload, two application parameters and three system parameters.
	\end{umonsThesisAbstract}

@@

	\chapter{Introduction}
	This chapter will...
	\include{intro/intro}

	\chapter{Application Models} \label{chapter:models}
	In this chapter, the proposed power, performance and energy models are derived and validated.
	Energy consumption is crucial in high-performance computing (HPC), especially to enable the next exascale generation. Hence, modern systems implement various hardware and software features for power management. Nonetheless, due to numerous different implementations, we can always push the limits of software to achieve the most efficient use of our hardware. To be
	energy efficient, the software relies on dynamic voltage and frequency scaling (DVFS), as well as dynamic power management (DPM). Yet, none have privileged information on the hardware architecture and application behavior, which may lead to energy-inefficient software operation. 
	This study  proposes analytical modeling for architecture and application behavior that can be used to estimate energy-optimal software configurations and provide knowledgeable hints to improve DVFS and DPM techniques for single-node HPC applications.
	Additionally, model parameters, such as the level of parallelism and dynamic power, provide insights into how the modeled application consumes energy, which can be helpful for energy-efficient software development and operation.
	This novel analytical model takes the number of active cores, the operating frequencies, and the input size as inputs to provide energy consumption estimation.
	We present the modeling of 13 parallel applications employed to determine energy-optimal configurations for several different input sizes.
	The results show that up to 70\% of energy could be saved in the best scenario compared to the default Linux choice and 14\% on average.
	We also compare the proposed model with standard machine-learning modeling concerning training overhead and accuracy. The results show that our approach generates about 10 times less energy overhead for the same level of accuracy.
	\include{models/models}
	%\include{optmizers/optmizers}

	\chapter{Energy optimization using the models}
	\include{experiments/setup}
	\include{optmization/optmization}

	\chapter{Extending the models with phases}
	\include{phases/phases}

	\chapter{Framework}
	This chapter introduces the framework developed, the fingerprint tool and pascal analyzer.

	\include{pascalanalyzer/pascalanalyzer}
	\include{fingerprint/fingerprint}

	This chapter will describe all the experiments that validate the models constraints.
	\include{framework/validation}

	\chapter{Conclusions}
	This chapter will...
	\include{conclusion}

>>>>

	\begin{umonsThesisAbstract}
		Energy consumption is a key to enabling exascale High performance Computing (HPC). However, energy-optimized hardware and software combinations could still be inefficient if the software operates poorly.

		Software operation relies on dynamic scaling of frequency and voltage (DVFS) and dynamic power management (DPM), but none have priority information on the application, leading to inefficient software operation.

		This work proposes a set of tools, models, and algorithms for energy optimization aimed at high-performance computing based on knowledge of the application and the specific architecture. The main contributions of this work are:

		A framework called PascalSuite automatically measures and compares multiple executions of a parallel application according to various scenarios characterized by input arrangements, number of threads, number of cores, and frequencies. As a result, PascalSuite can automate designing application models with an overhead of less than 1\%.

		A full-system energy model based on the CPU frequency and the number of cores. The model aims to understand and optimize the energy behavior of parallel applications in HPC systems according to application parameters, such as the degree of parallelism and CPU parameters related to dynamic and static power.

		A methodology that combines measurement data with a heuristic algorithm to provide insights into choosing the best phase divisions. Our heuristic can reduce the scan space from $10^{7000}$ to $10^2$ with an average error of 10\% and up to 38\% reduction in energy consumption using optimal distribution compared to standard Linux DVFS.
	\end{umonsThesisAbstract}

@@

	\chapter{Introduction} \label{chapter:introduction}
	In this chapter, we present the motivations for this work, as well as its objectives and main contributions.
	\include{intro/intro}

	\chapter{Theoretical background} \label{chapter:theoretical_background}
	This chapter will give the reader a better understanding of some terms used throughout the thesis, as well as details of configurations, the chosen benchmarks, and the architecture of the case study.
	\include{experiments/setup}

	\chapter{Parallel Scalability Suite} \label{chapter:pascal_suite}
	High-performance computing systems have become increasingly dynamic, complex, and unpredictable. To help build software that uses full-system capabilities, performance measurement and analysis tools exploit extensive execution analysis focusing on single-run results. Despite being effective in identifying performance hotspots and bottlenecks, these tools are not sufficiently suitable to evaluate the overall scalability trends of parallel applications. Either they lack the support for combining data from multiple runs or collect excessive data, causing unnecessary overhead. 
	In this chapter, we present a tool for automatically measuring and comparing several executions of a parallel application according to various scenarios characterized by the input arrangements, the number of threads, number of cores, and frequencies.
	Unlike other existing performance analysis tools, the proposed work covers some gaps in specialized features necessary to better understand computational resources scalability trends across configurations.
	In order to improve scalability analysis and productivity over the vast spectrum of possible configurations, the proposed tool features automatic instrumentation, direct mapping of parallel regions, accuracy-preserving data reductions, and ease of use.
	As it aims at accurately understanding scalability trends of parallel applications, detailed single-run performance analyses show minimal intrusion (less than 1\% overhead).

	\include{pascalanalyzer/pascalanalyzer}
	\include{fingerprint/fingerprint}
	\include{framework/validation}

	\chapter{Application energy and performance models} \label{chapter:models}
	Energy consumption is crucial in high-performance computing (HPC), especially to enable the next exascale generation. Hence, modern systems implement various hardware and software features for power management. Nonetheless, due to numerous different implementations, we can always push the limits of software to achieve the most efficient use of our hardware. To be energy efficient, the software relies on dynamic voltage and frequency scaling (DVFS), as well as dynamic power management (DPM). Yet, none have privileged information on the hardware architecture and application behavior, which may lead to energy-inefficient software operation. 
	This chapter proposes analytical modeling for architecture and application behavior that can be used to estimate energy-optimal software configurations and provide knowledgeable hints to improve DVFS and DPM techniques for single-node HPC applications.
	Additionally, model parameters, such as the level of parallelism and dynamic power, provide insights into how the modeled application consumes energy, which can be helpful for energy-efficient software development and operation.
	This novel analytical model takes the number of active cores, the operating frequencies, and the input size as inputs to provide energy consumption estimation.
	In this chapter we present the modeling of 13 parallel applications employed to determine energy-optimal configurations for several different input sizes.
	The results show that up to 70\% of energy could be saved in the best scenario compared to the default Linux choice and 14\% on average.
	We also compare the proposed model with standard machine-learning modeling concerning training overhead and accuracy. The results show that our approach generates about 10 times less energy overhead for the same level of accuracy.
	\include{models/models}
	%\include{optmizers/optmizers}
	\include{optmization/optmization}

	\chapter{Application-phase heuristic energy optimization} \label{chapter:phases}
	Dynamic voltage and frequency scaling (DVFS) and dynamic power management (DPM) are essential techniques for saving energy in modern computers. These techniques control hardware resources such as frequency and number of active processors. Although there is much work on the subject, there is an underexplored sub-topic: the impact of the choice of phase division on energy consumption.
	This chapter analyzes the impact of phase division on the energy efficiency of algorithms and proposes a methodology that combines measurement data with a heuristic to provide insights into choosing the best phase divisions.
	Our heuristic can reduce the scan space from $10^{7000}$ to $10^2$ with an average error of 10\% and up to 38\% reduction in energy consumption using optimal distribution compared to standard Linux DVFS. Moreover, we evaluated the trade-off of having too many divisions and the overhead caused. Finally, we have identified that there is usually a limit to the number of phases a running application can benefit from, giving a lower limit to the minimum number of phases.
	\include{phases/phases}

	\chapter{Conclusions and future work*} \label{chapter:conclusions}
	In this chapter, we deal with the conclusions about the effectiveness of the frameworks and the advantages and disadvantages of using models and algorithms with additional information on applications and architectures for energy optimization.
	We also discussed what could be improved in each approach and framework.
	\include{conclusion}

### 2022-08-15 thesis/pascalanalyzer/pascalanalyzer.tex
organazing the chapters and trying to connect

\section{PascalAnalyzer} \label{sec:pascal_analyzer}
High-performance computing systems have become increasingly dynamic, complex, and unpredictable. To help build software that uses full-system capabilities, performance measurement and analysis tools exploit extensive execution analysis focusing on single-run results. Despite being effective in identifying performance hotspots and bottlenecks, these tools are not sufficiently suitable to evaluate the overall scalability trends of parallel applications. Either they lack the support for combining data from multiple runs or collect excessive data, causing unnecessary overhead. 
In this work, we present a tool for automatically measuring and comparing several executions of a parallel application according to various scenarios characterized by the input arrangements, the number of threads, number of cores, and frequencies.
%In this work, we present a tool for measuring and comparing several executions of a parallel application automatically with different input arrangements, number of threads, cores, and frequencies to understand its scalability by observing how it uses the available computational resources across configurations. 
Unlike other existing performance analysis tools, the proposed work covers some gaps in specialized features necessary to better understand computational resources scalability trends across configurations. 
%The proposed work covers some gaps and unspecialized features of existing performance analysis tools. 
%The tool features automatic instrumentation and direct mapping of parallel regions, accuracy-preserving data reductions, and ease of use to compare multiple configurations for improving analysis and productivity. 
In order to improve scalability analysis and productivity over the vast spectrum of possible configurations, the proposed tool features automatic instrumentation, direct mapping of parallel regions, accuracy-preserving data reductions, and ease of use.
%In addition, it aims at parallel scalability, instead focus on single-run performance, aggregating a minimal level of intrusion (<1\% on average), which is fundamental for accurately understanding the behavior and scalability trends of the parallel application.
As it aims at accurately understanding scalability trends of parallel applications, detailed single-run performance analyses show minimal intrusion (less than 1\% overhead).

@@

Nowadays, due to the complexity of parallel systems, correctly identifying and locating performance and scalability bottlenecks depends on the developer's ability to compare several measurements in different execution configurations~\cite{Bergel2019, Silva2018}. This investigative process tends to be tedious and complex, requiring in-depth knowledge of the problem domain, parallel systems, and measurement and analysis tools. 
%Mastering the analysis tools becomes necessary mainly because they differ from each other in terms of measurement strategy, the collected metrics, and the many points and focus of observation. 
Mastering the analysis tools is necessary because they differ from each other in terms of measurement strategy, metrics collected, and the many points and focus of observation.
Even though some solutions propose organizing and combining metrics from different sources, these distinct characteristics can compel developers to exploit more than one tool depending on their analysis~interests.

%Despite their differences, every existing tool has its own contributions in helping to understand the inner working of a program, and some can present a vast and varied set of metrics with a high degree of detail. 
Every existing tool has its own contributions in helping to understand the inner working of a program. Some can present a vast and varied set of metrics with a high degree of detail. 
However, developers may have difficulty using these tools when their interest is in analyzing parallel scalability. 
%The main challenge arises from the focus on investigating a single run, because scalability analysis requires contrasting runs of different configurations. 
The main challenge arises from the emphasis on studying a single run because scalability analysis requires contrasting runs of different configurations.
%In this sense, it tends to be more significant to measure and compare only a few data across executions of multiple configurations than to collect a vast amount of finer-grain data of single-configuration runs.
In terms of productivity, measuring and comparing only a few data between runs of multiple configurations tends to be more significant to scalability analysis than collecting a vast amount of finer-grain data of single-configuration runs.
Furthermore, due to the number of measurements, there is a natural overhead associated with the execution of the analysis tool itself. Some works indicate this overhead can reach 40\% of the program runtime, which influences and damages the observation of its efficiency variation~\cite{Eriksson2016}.
As such, a measurement and analysis approach that uses only the data needed to understand the application's efficiency variation can be advantageous. First, because measuring and collecting only the necessary data limits the degree of intrusion and, second, because this strategy makes the analysis tool simpler and easier to use. 
%Finally, the tool's availability can also be a drawback as some apply to specific compilers and execution frameworks. Indeed, the framework version is a limitation for some traditional tools like HPCToolkit~\cite{Adhianto2010} and TAU (Tuning and Analysis Utilities)
%MDPI: Is this an abbreviation? If affirmative, please add the explanation. Please carefully check and confirm -- INCLUDED ON THE NEXT PARAGRAPH
~\cite{Shende2006}.

@@

The proposed tool uses a tracing-based measurement technique to identify and measure the boundaries of the analysis regions. The tracing approach supports a hierarchical region analysis model allowing users to see how the efficiency of inner parts can impact the program's overall scalability. In addition, the tool uses an aggregation feature to reduce the volume of data produced while preserving accurate analysis capability with minimal overhead. The tool also includes usability features that reduce the developer's efforts during the measurement and analysis process. These features allow one to automate the process of running and merging data collected from multiple configurations, bypassing manual efforts and avoiding the drawbacks of non-computer-oriented analysis. 
%The tool does not display visuals elements natively. Still, its interface allows the collected data to be exported and viewed from the control terminal in tabular form. Data exported via the interface are organized into data frames that can be interpreted by other tools or even rendered directly from graphic libraries.
The collected data, available from the terminal in tabular form, can be exported as data blocks that can be interpreted by other tools or even rendered by graphics libraries to facilitate the visualization and data analysis.

%The proposed tool is an alternative for realizing parallel scalability analysis more efficiently than single-run-centric performance measurement and analysis tools. 
We offer an alternative tool for realizing parallel scalability analysis more efficiently than single-run-centric performance measurement and analysis tools. In addition to that, our framework can also be used to promote energy-aware software; methods that rely on accurate and efficient energy collection could benefit from energy consumption data for the entire program and specific regions of the application \cite{10.1007/978-3-319-58667-0_22,10.1007/978-3-319-58667-0_21,7016382,10.1145/3321551}. To summarize, the main contributions of this work are:

>>>>

@@

Nowadays, due to the complexity of parallel systems, correctly identifying and locating performance and scalability bottlenecks depends on the developer's ability to compare several measurements in different execution configurations~\cite{Bergel2019, Silva2018}. This investigative process tends to be tedious and complex, requiring in-depth knowledge of the problem domain, parallel systems, and measurement and analysis tools.
Mastering the analysis tools is necessary because they differ from each other in terms of measurement strategy, metrics collected, and the many points and focus of observation.
Even though some solutions propose organizing and combining metrics from different sources, these distinct characteristics can compel developers to exploit more than one tool depending on their analysis~interests.

Every existing tool has its own contributions in helping to understand the inner working of a program. Some can present a vast and varied set of metrics with a high degree of detail. 
However, developers may have difficulty using these tools when their interest is in analyzing parallel scalability.
The main challenge arises from the emphasis on studying a single run because scalability analysis requires contrasting runs of different configurations.
In terms of productivity, measuring and comparing only a few data between runs of multiple configurations tends to be more significant to scalability analysis than collecting a vast amount of finer-grain data of single-configuration runs.
Furthermore, due to the number of measurements, there is a natural overhead associated with the execution of the analysis tool itself. Some works indicate this overhead can reach 40\% of the program runtime, which influences and damages the observation of its efficiency variation~\cite{Eriksson2016}.
As such, a measurement and analysis approach that uses only the data needed to understand the application's efficiency variation can be advantageous. First, because measuring and collecting only the necessary data limits the degree of intrusion and, second, because this strategy makes the analysis tool simpler and easier to use. 
~\cite{Shende2006}.

@@

The proposed tool uses a tracing-based measurement technique to identify and measure the boundaries of the analysis regions. The tracing approach supports a hierarchical region analysis model allowing users to see how the efficiency of inner parts can impact the program's overall scalability. In addition, the tool uses an aggregation feature to reduce the volume of data produced while preserving accurate analysis capability with minimal overhead. The tool also includes usability features that reduce the developer's efforts during the measurement and analysis process. These features allow one to automate the process of running and merging data collected from multiple configurations, bypassing manual efforts and avoiding the drawbacks of non-computer-oriented analysis.
The collected data, available from the terminal in tabular form, can be exported as data blocks that can be interpreted by other tools or even rendered by graphics libraries to facilitate the visualization and data analysis.

We offer an alternative tool for realizing parallel scalability analysis more efficiently than single-run-centric performance measurement and analysis tools. In addition to that, our framework can also be used to promote energy-aware software; methods that rely on accurate and efficient energy collection could benefit from energy consumption data for the entire program and specific regions of the application \cite{10.1007/978-3-319-58667-0_22,10.1007/978-3-319-58667-0_21,7016382,10.1145/3321551}. To summarize, the main contributions of this work are:
